---
title: "eGFR_values_AHEaD_simulated_data_2024"
output: html_document
date: "2024-07-15"
---

## Week 4

## Loading packages and data

```{r}
library(tidyverse)
library(here)
library(readr)

```

## eGFR Equations

***CKD-EPI Creatinine Equation (2009)***

[eGFRcr = 141 x min(Scr/κ, 1)α x max(Scr/κ, 1)-1.209 x 0.993Age x 1.018 [if female] x 1.159 [if black]]{.underline}

NOTE: What about for males? [if female], [if black]

where:

-   Scr = serum creatinine in mg/dL

-   κ = 0.7 (females) or 0.9 (males)

-   α = -0.329 (female) or -0.411 (male)

-   min(Scr/κ, 1) is the minimum of Scr/κ or 1.0

-   max(Scr/κ, 1) is the maximum of Scr/κ or 1.0

***CKD-EPI Creatinine Equation (2021)***

[eGFRcr = 142 x min(Scr/κ, 1)α x max(Scr/κ, 1)-1.200 x 0.9938Age x 1.012 [if female]]{.underline}

where:

-   Scr = standardized serum creatinine in mg/dL

-   κ = 0.7 (females) or 0.9 (males)

-   α = -0.241 (female) or -0.302 (male)

-   min(Scr/κ, 1) is the minimum of Scr/κ or 1.0

-   max(Scr/κ, 1) is the maximum of Scr/κ or 1.0

-   Age (years)

*Code Review from Sherri - Use the equation to assign/plug-in the variables rather than using the 'ifelse' function.*

## Variables that need to have eGFR computed based off their recorded values

1.  Serum creatinine (Sr_Cr)

2.  Age (age)

3.  Sex (sex)

4.  Race (Black or African American)

## Code

**CKD-EPI Creatinine Equation (2009)**

[eGFRcr = 141 x min(Scr/κ, 1)α x max(Scr/κ, 1)-1.209 x 0.993Age x 1.018 [if female] x 1.159 [if black]]{.underline}

-   κ = 0.7 (females) or 0.9 (males)

-   α = -0.329 (female) or -0.411 (male)

-   sex = female, male

-   "Black or African American" column = 1 or 0

```{r}
# Patient 1 (patient_id = 82842)
141*min(1.07/0.7, 1)^-0.329*max(1.07/0.7, 1)^-1.209*0.993^93*1.018*1.159

# Patient 2 (patient_id = 141877)
141*min(1.17/0.7, 1)^-0.329*max(1.17/0.7, 1)^-1.209*0.993^58*1.018*1.159

# Patient 3 (patient_id = 338629)
141*min(0.89/0.9, 1)^-0.329*max(0.89/0.9, 1)^-1.209*0.993^78*1.159

# Patient 4 (patient_id = 389931)
141*min(1.21/0.9, 1)^-0.329*max(1.21/0.9, 1)^-1.209*0.993^57*1.159

# Patient 5 (patient_id = 365017)
141*min(1.38/0.7, 1)^-0.329*max(1.38/0.7, 1)^-1.209*0.993^75*1.018*1.159
```

```{r}
### Generate new variable the assigns eGFR value to each participant using 2009 forumula
    ## Create a vector for the values of kappa that vary by binary sex
    k09 <- c("Female" = 0.7,
           "Male" = 0.9)
    
    ## Create a vector for the values of alpha that vary by binary sex
    a09 <- c("Female" = -0.329,
           "Male" = -0.411)
    
    ## Create a vector for the additional coefficient in the equation that varies by binary sex
    s09 <- c("Female" = 1.018,
           "Male" = 1)
    
    ## Create a vector for the additional coefficient in the equation that varies by race (Black vs. not)
    b09 <- c(1, 1.159) #R indexes starting at 1, so will need to shift (add 1) for Black indicator variable (assigns 1 to not Black and 1.159 to Black participants)
    
## Write out equation using above vectors
    eGFRcrD_2021_df$egfr09 <- 141 * 
      pmin(eGFRcrD_2021_df$Sr_Cr/k09[eGFRcrD_2021_df$sex], 1)^a09[eGFRcrD_2021_df$sex] *
      pmax(eGFRcrD_2021_df$Sr_Cr/k09[eGFRcrD_2021_df$sex], 1)^(-1.209) *
      (0.993^eGFRcrD_2021_df$age) * 
      (s09[eGFRcrD_2021_df$sex]) *
      (b09[1+eGFRcrD_2021_df$`Black or African American`])                   

    ## Filter to just Black participants to validate whether equation worked (against online calculator)
    testdf <- eGFRcrD_2021_df[eGFRcrD_2021_df$`Black or African American` >0,]
```

```{r}
#CKD09

#Stage 1 = 90>
#Stage 2 = >=60 + <90
#Stage 3a = >=45 +<60
#Stage 3b = >=30 + <45
#Stage 4 = >=15 + <30
#Stage 5 = <15

eGFRcrD_2021_df <- eGFRcrD_2021_df %>%  
  mutate(CKD09 = case_when(egfr09>=90~"Stage 1",
                           egfr09>=60 & egfr09<90~"Stage 2",
                           egfr09>=45 & egfr09<60~"Stage 3a",
                           egfr09>=30 & egfr09<45~"Stage 3b",
                           egfr09>=15 & egfr09<30~"Stage 4",
                           egfr09<15~"Stage 5"))
```

**CKD-EPI Creatinine Equation (2021)**

[eGFRcr = 142 x min(Scr/κ, 1)α x max(Scr/κ, 1)-1.200 x 0.9938Age x 1.012 [if female]]{.underline}

```{r}
# CKD-EPI Creatinine Equation (2021)

AHEaD_simulated_data_2024 <- read_csv("~/AHEaD_CKD_2024/AHEaD_simulated_data_2024/data/AHEaD_simulated_data_2024.csv")
View(AHEaD_simulated_data_2024)

# K
AHEaD_simulated_data_2024$K_21 <- ifelse(AHEaD_simulated_data_2024$sex=="Female",0.7,0.9)
table(AHEaD_simulated_data_2024$sex,AHEaD_simulated_data_2024$K_21)

# a
AHEaD_simulated_data_2024$a_21 <- ifelse(AHEaD_simulated_data_2024$sex=="Female", -0.241, -0.302)
table(AHEaD_simulated_data_2024$sex,AHEaD_simulated_data_2024$a_21)

# [if female]
AHEaD_simulated_data_2024$constant_21 <- ifelse(AHEaD_simulated_data_2024$sex=="Female",1.012,1)
table(AHEaD_simulated_data_2024$sex,AHEaD_simulated_data_2024$constant_21)

# Putting it altogether for the equation
AHEaD_simulated_data_2024$eGFRcr_21 <- 142*min(AHEaD_simulated_data_2024$Sr_Cr/AHEaD_simulated_data_2024$K_21, 1.0)*AHEaD_simulated_data_2024$a_21*max(AHEaD_simulated_data_2024$Sr_Cr/AHEaD_simulated_data_2024$K_21)*-1.200*0.9938*AHEaD_simulated_data_2024$age*AHEaD_simulated_data_2024$constant_21

# Histogram
hist(AHEaD_simulated_data_2024$eGFRcr_21)

# Summary
summary(AHEaD_simulated_data_2024$eGFRcr_21)

# Looking at the first 5 values
head(AHEaD_simulated_data_2024$eGFRcr_21, n=5)
```

```{r}
# Patient 1 (patient_id = 82842)
142*min(1.07/0.7, 1)^-0.241*max(1.07/0.7, 1)^-1.200*0.9938^93*1.012

# Patient 2 (patient_id = 141877)
141*min(1.17/0.7, 1)^-0.241*max(1.17/0.7, 1)^-1.200*0.9938^58*1.018*1.012

# Patient 3 (patient_id = 338629)
141*min(0.89/0.9, 1)^-0.302*max(0.89/0.9, 1)^-1.200*0.9938^78

# Patient 4 (patient_id = 389931)
141*min(1.21/0.9, 1)^-0.302*max(1.21/0.9, 1)^-1.200*0.9938^57

# Patient 5 (patient_id = 365017)
141*min(1.38/0.7, 1)^-0.241*max(1.38/0.7, 1)^-1.200*0.9938^75*1.018*1.012
```

```{r}
### Generate new variable the assigns eGFR value to each participant using 2021 forumula
    ## Create a vector for the values of kappa that vary by binary sex
    k21 <- c("Female" = 0.7,
           "Male" = 0.9)
    
    ## Create a vector for the values of alpha that vary by binary sex
    a21 <- c("Female" = -0.241,
           "Male" = -0.302)
    
    ## Create a vector for the additional coefficient in the equation that varies by binary sex
    s21 <- c("Female" = 1.012,
           "Male" = 1)
    
## Write out equation using above vectors
    eGFRcrD_2021_df$egfr21 <- 142 * 
      pmin(eGFRcrD_2021_df$Sr_Cr/k21[eGFRcrD_2021_df$sex], 1)^a21[eGFRcrD_2021_df$sex] *
      pmax(eGFRcrD_2021_df$Sr_Cr/k21[eGFRcrD_2021_df$sex], 1)^(-1.200) *
      (0.9938^eGFRcrD_2021_df$age) * 
      (s21[eGFRcrD_2021_df$sex])

    ## Filter to just Black participants to validate whether equation worked (against online calculator)
    testdf <- eGFRcrD_2021_df[eGFRcrD_2021_df$`Black or African American` >0,]
```

```{r}
#CKD21

#Stage 1 = 90>
#Stage 2 = >=60 + <90
#Stage 3a = >=45 +<60
#Stage 3b = >=30 + <45
#Stage 4 = >=15 + <30
#Stage 5 = <15

eGFRcrD_2021_df <- eGFRcrD_2021_df %>%
  mutate(CKD21 = case_when(egfr21>=90~"Stage 1",
                           egfr21>=60 & egfr21<90~"Stage 2",
                           egfr21>=45 & egfr21<60~"Stage 3a",
                           egfr21>=30 & egfr21<45~"Stage 3b",
                           egfr21>=15 & egfr21<30~"Stage 4",
                           egfr21<15~"Stage 5"))
```

## Week 5

```{r}
# Binary outcome for whether or not the patient has a different stage according to CKD-EPI 2009 and CKD-EPI 2021

eGFRcrD_2021_bo <- eGFRcrD_2021_df %>% 
  mutate(different_stage = ifelse(CKD09 != CKD21, 1, 0))

binary_outcome_summary <- eGFRcrD_2021_df %>%
  summarise(total_patients = n(),
    different_stage_count = sum(different_stage),
    same_stage_count = total_patients - different_stage_count,
    different_stage_percentage = (different_stage_count / total_patients) * 100,
  )

print(binary_outcome_summary)
```

```{r}
# Percent change between the CKD-EPI 2009 and CKD-EPI 2021 values

eGFRcrD_2021_pc <- eGFRcrD_2021_df %>% 
  mutate(percent_change = ((egfr21 - egfr09)/ egfr09)*100)

percent_change_summary <- eGFRcrD_2021_df %>%
  summarise(
    mean_percent_change = mean(percent_change, na.rm = TRUE),
    median_percent_change = median(percent_change, na.rm = TRUE),
    sd_percent_change = sd(percent_change, na.rm = TRUE))

print(percent_change_summary)
```

```{r}
# Assuming patient characteristics are in the dataset
# Example variables: age, sex, race, risk factors...

# Convert sex to a factor if they are not already
eGFRcrD_2021_df <- eGFRcrD_2021_df %>%
  mutate(sex = as.factor(sex))

# Logistic regression model (binomial - 2 responses)
logistic_model <- glm(different_stage ~ age + sex + Asian + Addl + Unknown + `Native Hawaiian or Addl Pacific Islander` + `Black or African American` + `American Indian or Alaska Native` + white + hypertension + diabetes + CKD + ESRD, 
                      data = eGFRcrD_2021_df, 
                      family = binomial)

# Summary of the logistic regression model
summary(logistic_model)

# Calculate odds ratios and 95% confidence intervals
exp(cbind(Odds_Ratio = coef(logistic_model), confint(logistic_model)))
```

```{r}
#Catherine's code - Black or African American
model_1 <- glm(different_stage ~ age + sex + `Black or African American` + hypertension + diabetes, 
                      data = eGFRcrD_2021_df, 
                      family = binomial)

exp(coef(model_1))

exp(cbind(Odds_Ratio = coef(model_1), confint(model_1)))
```

```{r}
#American Indian or Alaska Native
model_2 <- glm(different_stage ~ age + sex + `American Indian or Alaska Native` + hypertension + diabetes, 
                      data = eGFRcrD_2021_df, 
                      family = binomial)

exp(cbind(Odds_Ratio = coef(model_2), confint(model_2)))
```

```{r}
#White
model_3 <- glm(different_stage ~ age + sex + white + hypertension + diabetes, 
                      data = eGFRcrD_2021_df, 
                      family = binomial)

exp(cbind(Odds_Ratio = coef(model_3), confint(model_3)))
```

```{r}
# Reclassified patients

binary_outcome_summary <- eGFRcrD_2021_df %>% 
  summarise(total_patients = n(),
    different_stage_count = sum(different_stage),
    same_stage_count = total_patients - different_stage_count,
    different_stage_percentage = (different_stage_count / total_patients) * 100,
    same_stage_percentage = (same_stage_count / total_patients) * 100
  )

print(binary_outcome_summary)
```
